Assignment 2 :
========================================================

The main objective of the assignment is to find the relationship between the interest rate and the loan variables considering the Fico score.

Action Plan 
--------------------------------------------

1. First do exploratory analysis
2. Make some exploratory plots where you focus on the other variables . You can probably then divide those into 2 groups : those with little influence and those that have some. 
3. At that point you can do some more detailed investigation on the latter group. 
4. You might then do some formal tests on their correlation, both with the interest rate and with each other (confounders).
5. Then perhaps see if you can build a model where FICO and your selected variables can be used to predict interest rates. Does that tell you anything ? 

6. Then write up your methods, findings and results, paying close attention to the rubric (marking script). 

PS . One other point - it may be useful to have FICO both as categorical (binned) variable and as a continuous numeric variable. Plenty of other threads on this and how to do it if you wish.


When you click the **Knit HTML** button a web page will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r}
loan <-read.csv("loan.csv")
summary(loan)
```


Action Plan Step 1: Exploratory Analysis
--------------------------------------------

Purpose : To find out the relationship between the interest rate and other variables  and their significance.
Action Plan : 2. Make some exploratory plots where you focus on the other variables . You can probably then divide those into 2 groups : those with little influence and those that have some. 

 ****Open Credit Lines vs Interest rate****
 
 Inspect data distribution for open credit lines
 ```{r fig.width=10, fig.height=6}
 par(mfrow=c(1,2))
 hist(loan$open.credit.lines)
 plot(loan$interest.rate ~ loan$open.credit.lines)
 ```
 
 Data inspection difficult, hence we take the log. Further log is also used to transform the skewed data distribution to normal data distribution.
 ```{r fig.width=10, fig.height=6}
 par(mfrow=c(1,2))
 hist(log10(loan$open.credit.lines+1))
 plot(loan$interest.rate ~ log10(loan$open.credit.lines+1)) 
 ````
 Conclusion : **NO RELN OBSERVED** No pattern is observed during visual inspection. Hence the varaible open credit lines is concluded to be the variable with very slight relation with the interest rate.
 
 ****VARIABLE : Revolving Credit Balance vs Interest rate****
 
 Inspect data distribution for open credit lines
 ```{r fig.width=10, fig.height=6}
 par(mfrow=c(1,2))
 hist(loan$revolving.credit.balance)
 plot(loan$interest.rate ~ loan$revolving.credit.balance)
 ```
 
 Data inspection difficult, hence we take the log. Further log is also used to transform the skewed data distribution to normal data distribution.
 
 ````{r fig.width=10, fig.height=6}
 par(mfrow=c(1,2))
 hist(log10(loan$revolving.credit.balance+1))
 plot(loan$interest.rate ~ log10(loan$revolving.credit.balance+1))
 ````
 Visual inspection too difficult, so we plot linear model line to better see the reln
 
 ````{r}
 plot(loan$interest.rate ~ log10(loan$revolving.credit.balance+1))
 lm <- lm(loan$interest.rate ~ log10(loan$revolving.credit.balance+1))
 lines(log10(loan$revolving.credit.balance+1), lm$fitted, col = "red", lwd = 3)
 ````
 
 Conclusion : **OKAY**, Shows slightly more impact, hence considered for detail analysis for later part.
 
 
 ****VARIABLE : Inquiries in the last 6 month vs Interest rate****
 
 Inspect data distribution for open credit lines
 
 ```{r fig.width=10, fig.height=6}
 par(mfrow=c(1,2))
 hist(loan$inquiries.in.the.last.6.months)
 plot(loan$interest.rate ~ loan$inquiries.in.the.last.6.months)
 ```
 
 Data inspection difficult, hence take log
 
 ```{r fig.width=10, fig.height=6}
 par(mfrow=c(1,2))
 hist(log10(loan$inquiries.in.the.last.6.months+1))
 plot(loan$interest.rate ~ log10(loan$inquiries.in.the.last.6.months+1))
 ```
 
 Visual inspection too difficult, so we plot linear model line to better see the reln
 
 ````{r}
 par(mfrow=c(1,2))
 hist(log10(loan$inquiries.in.the.last.6.months+1))
 lm <- lm(loan$interest.rate ~ log10(loan$inquiries.in.the.last.6.months+1))
 plot(loan$interest.rate ~ log10(loan$inquiries.in.the.last.6.months+1))
 lines(log10(loan$inquiries.in.the.last.6.months+1), lm$fitted, col = "red", lwd = 3)
 ```
 
 Conclusion : **OKAY**, Shows slightly more impact, hence considered for detail analysis for later part.
 
***Attach the loan data to the data frame so that the columsn can then be acessed directly ***
dont forget to detach later using detach(loan)
```{r}
attach(loan)
```


****VARIABLE : Employement length vs Interest rate****
 
**Reformating of the data**
we can convert the emp lenth to ordered factor  and then treat or plot  the varaible as numeric
or we can  convert the pordered employement length to numeric.

```{r}
employment.length <- as.numeric(employment.length)
employment.length[is.na(employment.length)] <- 0
table(employment.length)
```


Inspect data distribution for open credit lines
 
 ```{r fig.width=10, fig.height=6}
 par(mfrow=c(1,2))
 hist(employment.length)
 plot(interest.rate ~ employment.length)
 lm_emplen <- lm(interest.rate ~ employment.length)
 lines(employment.length, lm_emplen$fitted, col = "red")
 ```


 Conclusion : **NOT RELEVANT**, Shows almost no impact on the interest rate, when the line was plotted.
 
**1. Amount.requested vs inerest.rate**

Inspect data distribution for open credit lines
 
 ```{r fig.width=10, fig.height=6}
 par(mfrow=c(1,2))
 hist(amount.requested)
 plot(interest.rate ~ amount.requested)
 lm_amtreq <- lm(interest.rate ~ amount.requested)
 lines(amount.requested, lm_amtreq$fitted, col = "red")
 lm_amtreq
 summary(lm_amtreq)
 ```
**1. fico.range vs inerest.rate**

Inspect data and plot
 
 ```{r fig.width=10, fig.height=6} 
 par(mfrow=c(1,2))
 range <- as.numeric(fico.range)
 hist(range)
 plot(interest.rate ~ range)
 lm_amtreq <- lm(interest.rate ~ range)
 lines(range, lm_amtreq$fitted, col = "red")
 lm_amtreq
 summary(lm_amtreq)
 ```
convert fico range to numeric with  fico.low value

```{r}
fico.low <-  as.numeric(lapply(strsplit(as.character(fico.range),"-"),"[",1))
loan$fico.low <- fico.low
hist(fico.low)
 plot(interest.rate ~ fico.low)
 lm_amtreq <- lm(interest.rate ~ fico.low)
 lines(fico.low, lm_amtreq$fitted, col = "red")
 lm_amtreq
 summary(lm_amtreq)

```


Plot fico as categrorical val
 
 ```{r fig.width=10, fig.height=6}
 par(mfrow=c(1,2))  
 plot(interest.rate ~ fico.range, varwidth = T)
 lm_amtreq <- lm(interest.rate ~ fico.range) 
 lm_amtreq
 summary(lm_amtreq)
 ```


ACtion STEP 2: Divide those into 2 groups : those with little influence and those that have some. 
----------------------------------------------------------------------------------

                                  T val        slope       p
1. amount,requested -------------- 14.85 -------- 0.00017 -------- 2e-16 \n
2. amoiunt.funde.by.investors -----17.95 --------0.00018 -------- "
3. loan.length --------------------23.34 -------- 4.27204 -------- <2e-16 
4. loan.purpose  -------- -------- yes important, qualitative data
5. debt.to.income.ration ----------8.6 -------- 0.095 -------- "
6. state --------  --------  ------Fval = 1.21
7. home.ownership --------  ------Fval = 5.87 yes
8. open.credit.lines --------  no reln observed during exploratory graph
9. revolving.credit.balance  -------- go for further exploration 
10.----------------------------3.13    0.000015   0.0018 (**REjected** , cause low T val, with very very low slope val and relatively high p val)
11. inquiries in the last.6.moths further explore
12.                            8.398   0.56332          <2e-16 
13. employement.length          no

REordering by tval 
                            T val        slope       p
1. loan.length                 23.34     4.27204    <2e-16 
2. amoiunt.funde.by.investors  17.95     0.00018     "
3. amount,requested            14.85     0.00017    2e-16
4. inquiries in the last.6.moths further explore
5.                             8.398   0.56332          <2e-16 
6. debt.to.income.ration       8.6       0.095       "
7. loan.purpose                yes important, qualitative data
8. home.ownership              Fval = 5.87 yes
9. state                       Fval = 1.21

```{r}
#Rejected factors
#
#1. open.credit.lines           no reln observed during exploratory graph
#2. revolving.credit.balance    go for further exploration 
#3.                             3.13    0.000015   0.0018 (**REjected** , cause low T val, with very very low 4. #slope val and relatively high p val)
#4. employement.length          no

```


*** Further analyse few variables ***

***Revolving.credit.balance vs interest rate***

  ```{r fig.width=10, fig.height=6}
 par(mfrow=c(1,2))
 hist(log10(revolving.credit.balance))
 plot(interest.rate ~ log10(revolving.credit.balance+1))
 lm_amtreq <- lm(interest.rate ~ log10(revolving.credit.balance+1))
 lines(log10(revolving.credit.balance+1), lm_amtreq$fitted, col = "red")
 lm_amtreq
 summary(lm_amtreq)
 ```

without log, distribution

 ```{r fig.width=10, fig.height=6}
 par(mfrow=c(1,2))
 hist(revolving.credit.balance)
 plot(interest.rate ~ revolving.credit.balance)
 lm_amtreq <- lm(interest.rate ~ revolving.credit.balance)
 lines(revolving.credit.balance, lm_amtreq$fitted, col = "red")
 lm_amtreq
 summary(lm_amtreq)
 ```

***Inquiries.in.the.last.6.months vs interest rate***


without log, distribution

 ```{r fig.width=10, fig.height=6}
 par(mfrow=c(1,2))
 hist(inquiries.in.the.last.6.months)
 plot(interest.rate ~ inquiries.in.the.last.6.months)
 lm_amtreq <- lm(interest.rate ~ inquiries.in.the.last.6.months)
 lines(inquiries.in.the.last.6.months, lm_amtreq$fitted, col = "red")
 lm_amtreq
 summary(lm_amtreq)
 ```


ACTION STEP 4 : do some formal tests on their correlation, both with the interest rate and with each other (confounders).
-------------------------------------------------------------------

calculating correlation across all numerical values. We should remove variable if they are too much co related. It implies that the bariables are possible similar
```{r  fig.width=15}

numeric.columns <- names(loan)[unlist(lapply(loan,"class")) %in% c("numeric", "integer")]
dn <- as.matrix(loan[,numeric.columns])
cordn <- cor( dn[complete.cases(dn),] )  
cordn
```

Conclusion :Observed Confounding varaible, amount requested and amount funded.
strong correlation, almost perfect correlation betn amount requested and amount funded. hence during analysis, we would include amount_requested. This makes mer sense logically as well.



ACTION STEP 5 : Build a model where FICO and your selected variables can be used to predict interest rates. Does that tell you anything ? 
---------------------------------------------------------
variables we would inspect
1. fico.low
2. loan.length                 23.34     4.27204    <2e-16 
3. amount,requested            14.85     0.00017    2e-16
4. inquiries in the last.6.moths further explore
5.                           8.398   0.56332          <2e-16 
6. debt.to.income.ration       8.6       0.095       "
7. loan.purpose                yes important, qualitative data
8. home.ownership              Fval = 5.87 yes
9. state                       Fval = 1.21

**building linear model with  interest rate, fico.low score and loan.length**

```{r}
attach(loan)
lm1 <- lm(interest.rate ~ fico.low + loan.length)
par(mfrow = c(1, 2))
plot(fico.low, lm1$residuals, pch = 19, col = loan.length)
plot(lm1$fitted, lm1$residuals, pch = 19, col = loan.length)
lm1
summary(lm1)
```

**building linear model with  interest rate, fico.low score and loan.length**

```{r}
attach(loan)
lm2 <- lm(interest.rate ~  loan.length + fico.low )
par(mfrow = c(1, 2))
plot(fico.low, lm2$residuals, pch = 19, col = loan.length)
plot(lm2$fitted, lm2$residuals, pch = 19, col = loan.length)
lm2
summary(lm2)
```




6. Then write up your methods, findings and results, paying close attention to the rubric (marking script). 

PS . One other point - it may be useful to have FICO both as categorical (binned) variable and as a continuous numeric variable. Plenty of other threads on this and how to do it if you wish.
